---
title: "TimeSeries2"
author: "Stacey Jackson"
date: "30/04/2019"
output: html_document
---
####1. Open libraries
```{r}

pacman::p_load(RMySQL, dplyr, lubridate, ggplot2,ggfortify,plotly, forecast,zoo, padr,rstudioapi,stats)

```

```{r}
con = dbConnect(MySQL(), user='deepAnalytics', password='Sqltask1234!', dbname='dataanalytics2018', host='data-analytics-2018.cbrosir2cswx.us-east-1.rds.amazonaws.com')
```

####2. SORT OUT THE DATASET
Download date, time and the sub-metering attributes for each sub-meter
```{r}
yr_2006SELECT <- dbGetQuery(con, "SELECT Date, Time, Global_active_power, Sub_metering_1, Sub_metering_2, Sub_metering_3 FROM yr_2006")
yr_2007SELECT <- dbGetQuery(con, "SELECT Date, Time, Global_active_power, Sub_metering_1, Sub_metering_2, Sub_metering_3 FROM yr_2007")
yr_2008SELECT <- dbGetQuery(con, "SELECT Date, Time, Global_active_power, Sub_metering_1, Sub_metering_2, Sub_metering_3 FROM yr_2008")
yr_2009SELECT <- dbGetQuery(con, "SELECT Date, Time, Global_active_power, Sub_metering_1, Sub_metering_2, Sub_metering_3 FROM yr_2009")
yr_2010SELECT <- dbGetQuery(con, "SELECT Date, Time, Global_active_power, Sub_metering_1, Sub_metering_2, Sub_metering_3 FROM yr_2010")
```
CREATE ONE DATAFRAME - ONLY YEARS THAT ARE COMPLETE
```{r}
newDF2 <- bind_rows(yr_2007SELECT, yr_2008SELECT, yr_2009SELECT,yr_2010SELECT)
```
Combine Date & Time attribute values in a new attribute column
```{r}
newDF2 <-cbind(newDF2,paste(newDF2$Date,newDF2$Time), stringsAsFactors=FALSE)
```
Give the new attribute in the 7th column a header name 
```{r}
colnames(newDF2)[7] <-"DateTime"
```
Move the DateTime attribute within the dataset
```{r}
newDF2 <- newDF2[,c(ncol(newDF2), 1:(ncol(newDF2)-1))]
```
Lubridate
```{r}
newDF2$DateTime <- ymd_hms(newDF2$DateTime)
newDF2$Time<-hms(newDF2$Time)
newDF2$Date<-ymd(newDF2$Date)
newDF2$year <- year(newDF2$DateTime)
newDF2$week<-week(newDF2$DateTime)
newDF2$weekdays<-wday(newDF2$DateTime)
newDF2$day <- day(newDF2$DateTime)
newDF2$hour <- hour(newDF2$DateTime)
newDF2$minute <- minute(newDF2$DateTime)
newDF2$quarter <- quarter(newDF2$DateTime)
newDF2$month<-month(newDF2$DateTime)

str(newDF2)
head(newDF2)

```
#### 3. MISSING VALUES
```{r}

#create empty rows
newDF2 <- pad(newDF2,by="DateTime",break_above=3)
head(newDF2)

sum(is.na(newDF2))


NA_mean <- replace(newDF2,TRUE, lapply(newDF2,na.aggregate))

sum(is.na(NA_mean))

head(NA_mean)##TIME HAS DISAPPEARED
str(NA_mean)
```


#### 3. Group by month
```{r}

house_month <- NA_mean%>%
  group_by(year,month)%>%
  summarize(month_total=sum(Global_active_power))

ts_month<-ts(house_month$month_total, frequency=12,start=c(2007,1))
ts_month

ggplot2::autoplot(ts_month)



# Decompose into trend, seasonal and remainder
decomposed_ts_month <- stl(ts_month, s.window = "periodic")
decomposed_ts_month

ggplot2::autoplot(decomposed_ts_month)


## Check variance
variance_ts_month<-apply(decomposed_ts_month$time.series,2,var)/var(ts_month)
variance_ts_month
                         
#create training and test sets

train_month<-window (ts_month,start=c(2007,1), end=c(2010,1))

test_month <- window(ts_month,start=c(2010,2))

##create a Holt Winters model
HW_month <- HoltWinters(train_month)

#check residuals for the model
autoplot(HW_month)
checkresiduals(HW_month)

#make forecasts
forecast_HW_month <- forecast:::forecast.HoltWinters(HW_month,h=9)

#plot time series with the forecasts
#ggplot2::autoplot(ts_month) + 
 # ggplot2::autolayer(forecast_HW_month$mean)

#check accuracy
HW_accuracy_month<-accuracy(forecast_HW_month,test_month)

##create an ARIMA model
AR_month<-auto.arima(train_month)

#check residuals for the model
ggplot2::autoplot(AR_month)
checkresiduals(AR_month)

##make forecasts
forecast_AR_month<-forecast:::forecast(AR_month,h=9)

#plot time series with the forecasts
#ggplot2::autoplot(ts_month) +
 # ggplot2::autolayer(forecast_AR_month$mean)

#check accuracy
AR_accuracy_month<-accuracy(forecast_AR_month,test_month)
AR_accuracy_month

```
#### 4. Group by week
```{r}
house_week <- NA_mean%>%
  group_by(year,week)%>%
  summarize(week_total=sum(Global_active_power))

ts_week<-ts(house_week$week_total, frequency=52,start=c(2007,1))

ggplot2::autoplot(ts_week)


## Decompose into trend, seasonal and remainder
decomposed_ts_week <- stl(ts_week, s.window = "periodic")
decomposed_ts_week

ggplot2::autoplot(decomposed_ts_month)


## Check variance
variance_ts_week<-apply(decomposed_ts_week$time.series,2,var)/var(ts_week)
variance_ts_week
##create training and test sets

train_week<-window (ts_week,start=c(2007,1), end=c(2010,1))

test_week <- window(ts_week,start=c(2010,2))

##create a Holt Winters model
HW_week <- HoltWinters(train_week)

##make forecasts
forecast_week <- forecast:::forecast.HoltWinters(HW_week,h=50)

#check residuals for the model
ggplot2::autoplot(HW_week)
checkresiduals(HW_week)

#plot time series with the forecasts
#ggplot2::autoplot(ts_week) +
 # ggplot2::autolayer(forecast_week$mean)

#check accuracy
HW_accuracy_week<-accuracy(forecast_week,test_week)
HW_accuracy_week

##create an ARIMA model
AR_week<-auto.arima(train_week)

#check residuals for the model
ggplot2::autoplot(AR_week)
checkresiduals(AR_week)

##make forecasts
forecast_AR_week<-forecast:::forecast(AR_week,h=50)

#plot time series with the forecasts
#ggplot2::autoplot(ts_week) +
 # ggplot2::autolayer(forecast_AR_week$mean)

#check accuracy
AR_accuracy_week<-accuracy(forecast_AR_week,test_week)
AR_accuracy_week
```
#### 4. Group by day
```{r}
house_day <- NA_mean%>%
  group_by(year,day)%>%
  summarize(day_total=sum(Global_active_power))

ts_day<-ts(house_day$day_total, frequency=365,start=c(2007,1))

ggplot2::autoplot(ts_day)


## Decompose into trend, seasonal and remainder
decomposed_ts_day <- stl(ts_day, s.window = "periodic")

ggplot2::autoplot(decomposed_ts_day)


## Check variance??????
variance_ts_day<-apply(decomposed_ts_day$time.series,2,var)/var(ts_day)
variance_ts_day
##create training and test sets

train_day<-window (ts_day,start=c(2007,1), end=c(2010,1))

test_day <- window(ts_day,start=c(2010,2))

##create a Holt Winters model
HW_day <- HoltWinters(train_day)

##make forecasts
forecast_day <- forecast:::forecast.HoltWinters(HW_day,h=363)

#check residuals for the model
ggplot2::autoplot(HW_day)

#plot time series with the forecasts
#ggplot2::autoplot(ts_day) +
 # ggplot2::autolayer(forecast_day$mean)

#check accuracy
HW_accuracy_day<-accuracy(forecast_day,test_day)
```
By quarter???? WHAT TIME PERIOD FOR THE FORECAST????
```{r}
house_quarter <- NA_mean%>%
  group_by(year,quarter)%>%
  summarize(month_quarter=sum(Global_active_power))

ts_quarter<-ts(house_quarter$month_quarter, frequency=4,start=c(2007,1))
ts_quarter

ggplot2::autoplot(ts_quarter)



# Decompose into trend, seasonal and remainder
decomposed_ts_quarter <- stl(ts_quarter, s.window = "periodic")
decomposed_ts_quarter

ggplot2::autoplot(decomposed_ts_quarter)


## Check variance
variance_ts_quarter<-apply(decomposed_ts_quarter$time.series,2,var)/var(ts_quarter)
variance_ts_quarter
                         
#create training and test sets

train_quarter<-window (ts_quarter,start=c(2007,1), end=c(2010,1))

test_quarter <- window(ts_quarter,start=c(2010,2))

##create a Holt Winters model
HW_quarter <- HoltWinters(train_quarter)

#check residuals for the model
autoplot(HW_quarter)
checkresiduals(HW_quarter)

#make forecasts
forecast_HW_quarter <- forecast:::forecast.HoltWinters(HW_quarter,h=9)

#plot time series with the forecasts
#ggplot2::autoplot(ts_quarter) + 
 # ggplot2::autolayer(forecast_HW_quarter$mean)

#check accuracy
HW_accuracy_quarter<-accuracy(forecast_HW_quarter,test_quarter)

##create an ARIMA model
AR_month<-auto.arima(train_quarter)

#check residuals for the model
ggplot2::autoplot(AR_quarter)
checkresiduals(AR_quarter)

##make forecasts
forecast_AR_quarter<-forecast:::forecast(AR_quarter,h=9)

#plot time series with the forecasts
#ggplot2::autoplot(ts_quarter) +
 # ggplot2::autolayer(forecast_AR_quarter$mean)

#check accuracy
AR_accuracy_quarter<-accuracy(forecast_AR_quarter,test_quarter)
AR_accuracy_quarter

```
```

