---
title: "TimeSeries"
author: "Stacey Jackson"
date: "24/04/2019"
output:
  html_document: default
  word_document: default
---
####1. Open libraries
```{r}

library(RMySQL)

library(dplyr)

library(lubridate)

library(ggplot2)

library(ggfortify)

library(tidyr)

library(plotly)

library(forecast)


```


```{r}
con = dbConnect(MySQL(), user='deepAnalytics', password='Sqltask1234!', dbname='dataanalytics2018', host='data-analytics-2018.cbrosir2cswx.us-east-1.rds.amazonaws.com')
```

####2. SORT OUT THE DATASET
Download date, time and the sub-metering attributes for each sub-meter
```{r}
yr_2006SELECT <- dbGetQuery(con, "SELECT Date, Time, Global_active_power, Sub_metering_1, Sub_metering_2, Sub_metering_3 FROM yr_2006")
yr_2007SELECT <- dbGetQuery(con, "SELECT Date, Time, Global_active_power, Sub_metering_1, Sub_metering_2, Sub_metering_3 FROM yr_2007")
yr_2008SELECT <- dbGetQuery(con, "SELECT Date, Time, Global_active_power, Sub_metering_1, Sub_metering_2, Sub_metering_3 FROM yr_2008")
yr_2009SELECT <- dbGetQuery(con, "SELECT Date, Time, Global_active_power, Sub_metering_1, Sub_metering_2, Sub_metering_3 FROM yr_2009")
yr_2010SELECT <- dbGetQuery(con, "SELECT Date, Time, Global_active_power, Sub_metering_1, Sub_metering_2, Sub_metering_3 FROM yr_2010")
```
CREATE ONE DATAFRAME - ONLY YEARS THAT ARE COMPLETE
```{r}
newDF2 <- bind_rows(yr_2007SELECT, yr_2008SELECT, yr_2009SELECT)
```
Combine Date & Time attribute values in a new attribute column
```{r}
newDF2 <-cbind(newDF2,paste(newDF2$Date,newDF2$Time), stringsAsFactors=FALSE)
```
Give the new attribute in the 7th column a header name 
```{r}
colnames(newDF2)[7] <-"DateTime"
```
Move the DateTime attribute within the dataset
```{r}
newDF2 <- newDF2[,c(ncol(newDF2), 1:(ncol(newDF2)-1))]
```
Lubridate
```{r}
newDF2$DateTime <- ymd_hms(newDF2$DateTime)
newDF2$Time<-hms(newDF2$Time)
newDF2$Date<-ymd(newDF2$Date)
newDF2$year <- year(newDF2$DateTime)
newDF2$week<-week(newDF2$DateTime)
newDF2$weekdays<-wday(newDF2$DateTime)
newDF2$day <- day(newDF2$DateTime)
newDF2$hour <- hour(newDF2$DateTime)
newDF2$minute <- minute(newDF2$DateTime)
newDF2$quarter <- quarter(newDF2$DateTime)
newDF2$month<-month(newDF2$DateTime)

str(newDF2)
head(newDF2)

```

#### 3. GRANULARITY
####3.1 Plot all of sub-meter 1
```{r}
plot(newDF2$Sub_metering_1)
```
###3.2. Subset the second week of 2008 - All Observations
```{r}
houseWeek <- filter(newDF2, year == 2008 & week == 2)
```
###3.3. Plot subset houseWeek
```{r}
plot(houseWeek$Sub_metering_1)
```
####3.4 Subset the 9th day of January 2008 - All observations
```{r}
houseDay <- filter(newDF2, year == 2008 & month == 1 & day == 9)
```
#### 3.5 Plot sub-meter 1
```{r}
plot_ly(houseDay, x = ~houseDay$DateTime, y = ~houseDay$Sub_metering_1, type = 'scatter', mode = 'lines')
```
####3.6 Plot sub-meter 1, 2 and 3 with title, legend and labels - All observations 
```{r}
plot_ly(houseDay, x = ~houseDay$DateTime, y = ~houseDay$Sub_metering_1, name = 'Kitchen', type = 'scatter', mode = 'lines') %>%
 add_trace(y = ~houseDay$Sub_metering_2, name = 'Laundry Room', mode = 'lines') %>%
 add_trace(y = ~houseDay$Sub_metering_3, name = 'Water Heater & AC', mode = 'lines') %>%
 layout(title = "Power Consumption January 9th, 2008",
 xaxis = list(title = "Time"),
 yaxis = list (title = "Power (watt-hours)"))
```

####3.7 Subset the 9th day of January 2008 - 10 Minute frequency
```{r}
houseDay10 <- filter(newDF2, year == 2008 & month == 1 & day == 9 & (minute == 0 | minute == 10 | minute == 20 | minute == 30 | minute == 40 | minute == 50))
```
####3.8 Plot sub-meter 1, 2 and 3 with title, legend and labels - 10 Minute frequency
```{r}
plot_ly(houseDay10, x = ~houseDay10$DateTime, y = ~houseDay10$Sub_metering_1, name = 'Kitchen', type = 'scatter', mode = 'lines') %>%
 add_trace(y = ~houseDay10$Sub_metering_2, name = 'Laundry Room', mode = 'lines') %>%
 add_trace(y = ~houseDay10$Sub_metering_3, name = 'Water Heater & AC', mode = 'lines') %>%
 layout(title = "Power Consumption January 9th, 2008",
 xaxis = list(title = "Time"),
 yaxis = list (title = "Power (watt-hours)"))
```
####3.9 QUESTIONS
What peaks might represent the water heater? How about the AC? 
What could be happening in the laundry room? Only the fridge regulating temperature throughout the day in the laundry room.
How many times during this day are kitchen appliances being used? Kitchen appliances only used in the evening - used twice. 
Lastly, in your opinion, does the data from these three sub-meters contain useful information for the homeowner? 
Doesn't feel particularly useful.

###4.Subset the third week of 2008 - All Observations
```{r}
houseWeek <- filter(newDF2, year == 2008 & week == 3)
```
####4.1 Plot the third week
```{r}
plot_ly(houseWeek, x = ~houseWeek$DateTime, y = ~houseWeek$Sub_metering_1, name = 'Kitchen', type = 'scatter', mode = 'lines') %>%
 add_trace(y = ~houseWeek$Sub_metering_2, name = 'Laundry Room', mode = 'lines') %>%
 add_trace(y = ~houseWeek$Sub_metering_3, name = 'Water Heater & AC', mode = 'lines') %>%
 layout(title = "Power Consumption January 9th, 2008",
 xaxis = list(title = "Time"),
 yaxis = list (title = "Power (watt-hours)"))
```
####4.2 Subset the third week of January 2008 - 60 Minute frequency
```{r}
houseWeek60 <- filter(newDF2, year == 2008 & month == 1 & week == 3 & (minute == 0))
```
####4.3 Plot sub-meter 1, 2 and 3 with title, legend and labels - 60 Minute frequency
```{r}
plot_ly(houseWeek60, x = ~houseWeek60$DateTime, y = ~houseWeek60$Sub_metering_1, name = 'Kitchen', type = 'scatter', mode = 'lines') %>%
 add_trace(y = ~houseWeek60$Sub_metering_2, name = 'Laundry Room', mode = 'lines') %>%
 add_trace(y = ~houseWeek60$Sub_metering_3, name = 'Water Heater & AC', mode = 'lines') %>%
 layout(title = "Power Consumption Week Three, January , 2008",
 xaxis = list(title = "Time"),
 yaxis = list (title = "Power (watt-hours)"))
```
###5. Subset the first month of 2008 - All Observations
```{r}
houseMonth <- filter(newDF2, year == 2008 & month == 1)

plot_ly(houseMonth, x = ~houseMonth$DateTime, y = ~houseMonth$Sub_metering_1, name = 'Kitchen', type = 'scatter', mode = 'lines') %>%
 add_trace(y = ~houseMonth$Sub_metering_2, name = 'Laundry Room', mode = 'lines') %>%
 add_trace(y = ~houseMonth$Sub_metering_3, name = 'Water Heater & AC', mode = 'lines') %>%
 layout(title = "Power Consumption January, 2008",
 xaxis = list(title = "Time"),
 yaxis = list (title = "Power (watt-hours)"))
```
####5.1 Subset the first month of January 2008 - 6 hour frequency
```{r}
houseMonth6 <- filter(newDF2, year == 2008 & month == 1 & (hour == 0 | hour == 6 | hour==12 | hour==18))

plot_ly(houseMonth6, x = ~houseMonth6$DateTime, y = ~houseMonth6$Sub_metering_1, name = 'Kitchen', type = 'scatter', mode = 'lines') %>%
 add_trace(y = ~houseMonth6$Sub_metering_2, name = 'Laundry Room', mode = 'lines') %>%
 add_trace(y = ~houseMonth6$Sub_metering_3, name = 'Water Heater & AC', mode = 'lines') %>%
 layout(title = "Power Consumption January, 2008",
 xaxis = list(title = "Time"),
 yaxis = list (title = "Power (watt-hours)"))
```
####5.2 Subset the first month of January 2008 - 4 hour frequency
```{r}
houseMonth4 <- filter(newDF2, year == 2008 & month == 1 & (hour == 0 | hour == 4 | hour==12 | hour==16 | hour==20))

plot_ly(houseMonth4, x = ~houseMonth4$DateTime, y = ~houseMonth4$Sub_metering_1, name = 'Kitchen', type = 'scatter', mode = 'lines') %>%
 add_trace(y = ~houseMonth4$Sub_metering_2, name = 'Laundry Room', mode = 'lines') %>%
 add_trace(y = ~houseMonth4$Sub_metering_3, name = 'Water Heater & AC', mode = 'lines') %>%
 layout(title = "Power Consumption January, 2008",
 xaxis = list(title = "Time"),
 yaxis = list (title = "Power (watt-hours)"))
```

####5.3 Subset the first month of January 2008 - 3 hour frequency
```{r}
houseMonth3 <- filter(newDF2, year == 2008 & month == 1 & (hour == 0 | hour == 3 | hour==9 | hour==12 | hour==15 | hour==18 | hour== 21))

plot_ly(houseMonth3, x = ~houseMonth3$DateTime, y = ~houseMonth3$Sub_metering_1, name = 'Kitchen', type = 'scatter', mode = 'lines') %>%
 add_trace(y = ~houseMonth3$Sub_metering_2, name = 'Laundry Room', mode = 'lines') %>%
 add_trace(y = ~houseMonth3$Sub_metering_3, name = 'Water Heater & AC', mode = 'lines') %>%
 layout(title = "Power Consumption January, 2008",
 xaxis = list(title = "Time"),
 yaxis = list (title = "Power (watt-hours)"))
```
####5.4 Subset the first month of January 2008 - 8 hour frequency
```{r}
houseMonth8 <- filter(newDF2, year == 2008 & month == 1 & (hour == 0 | hour == 8 | hour==16))

plot_ly(houseMonth8, x = ~houseMonth8$DateTime, y = ~houseMonth8$Sub_metering_1, name = 'Kitchen', type = 'scatter', mode = 'lines') %>%
 add_trace(y = ~houseMonth8$Sub_metering_2, name = 'Laundry Room', mode = 'lines') %>%
 add_trace(y = ~houseMonth8$Sub_metering_3, name = 'Water Heater & AC', mode = 'lines') %>%
 layout(title = "Power Consumption January, 2008",
 xaxis = list(title = "Time"),
 yaxis = list (title = "Power (watt-hours)"))
```
####6.PIE CHARTS
####6.1 Pie chart - Total power use over 9 Jan 2008 by each sub-meter
```{r}

newDF2%>%
  filter(year==2008)%>%
  filter(month==1)%>%
  filter(day==9)%>%
  summarise(Sub_metering_1=sum(Sub_metering_1), Sub_metering_2=sum(Sub_metering_2), Sub_metering_3=sum(Sub_metering_3))%>%
  gather(key,value, Sub_metering_1, Sub_metering_2, Sub_metering_3)%>%
  plot_ly(labels=~key, values=~value, type='pie')%>%
  layout(title = 'Total power use over 9 Jan 2008 by each sub-meter',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

```
####6.2 Pie chart - Total power use on 9 Jan 2008 in the morning 7am-10am by each sub-meter and 5pm-10pm
```{r}

newDF2%>%
  filter(year==2008)%>%
  filter(month==1)%>%
  filter(day==9)%>%
  filter(hour>6 & hour<11)%>%
  summarise(Sub_metering_1=sum(Sub_metering_1), Sub_metering_2=sum(Sub_metering_2), Sub_metering_3=sum(Sub_metering_3))%>%
  gather(key,value, Sub_metering_1, Sub_metering_2, Sub_metering_3)%>%
  plot_ly(labels=~key, values=~value, type='pie')%>%
  layout(title = 'Total power use over 9 Jan 2008 by each sub-meter',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

newDF2%>%
  filter(year==2008)%>%
  filter(month==1)%>%
  filter(day==9)%>%
  filter(hour>16 & hour<23)%>%
  summarise(Sub_metering_1=sum(Sub_metering_1), Sub_metering_2=sum(Sub_metering_2), Sub_metering_3=sum(Sub_metering_3))%>%
  gather(key,value, Sub_metering_1, Sub_metering_2, Sub_metering_3)%>%
  plot_ly(labels=~key, values=~value, type='pie')%>%
  layout(title = 'Total power use over 9 Jan 2008 by each sub-meter',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

```

####6.3 Pie chart - Total power use over 2008 by each sub-meter
```{r}

newDF2%>%
  filter(year==2008)%>%
  summarise(Sub_metering_1=sum(Sub_metering_1), Sub_metering_2=sum(Sub_metering_2), Sub_metering_3=sum(Sub_metering_3))%>%
  gather(key,value, Sub_metering_1, Sub_metering_2, Sub_metering_3)%>%
  plot_ly(labels=~key, values=~value, type='pie')%>%
  layout(title = 'Total power use over 2008 by each sub-meter',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

```
####7. TIME SERIES ANALYSIS
####7.1 Create a Time Series object
```{r}
## Subset to one observation per week on Mondays at 8:00pm for 2007, 2008 and 2009
house070809weekly <- filter(newDF2, weekdays == 2 & hour == 20 & minute == 1)
```
#### 7.2 Create TS object with SubMeter3
```{r}
tsSM3_070809weekly <- ts(house070809weekly$Sub_metering_3, frequency=52, start=c(2007,1))
```
## 7.3 Plot sub-meter 3 with autoplot 
```{r}
autoplot(tsSM3_070809weekly)

## Plot sub-meter 3 with autoplot - add labels, color
autoplot(tsSM3_070809weekly, ts.colour = 'red', xlab = "Time", ylab = "Watt Hours", main = "Sub-meter 3")

## Plot sub-meter 3 with plot.ts
plot.ts(tsSM3_070809weekly)
```
####7.4 Sub-meter 1 - change to Saturdays because the kitchen and used more at the weekend - 7pm. Think they ate later one year!!!
####7.5 Create TS object with SubMeter1 - weekly
```{r}
house070809weekly <- filter(newDF2, weekdays == 7 & hour == 18 & minute == 1)

tsSM1_070809weekly <- ts(house070809weekly$Sub_metering_1, frequency=52, start=c(2007,1))
```
####7.6 Plot sub-meter 1 with autoplot - weekly
```{r}
autoplot(tsSM1_070809weekly)

## Plot sub-meter 2 with autoplot - add labels, color
autoplot(tsSM1_070809weekly, ts.colour = 'red', xlab = "Time", ylab = "Watt Hours", main = "Sub-meter 1")

## Plot sub-meter 1 with plot.ts
plot.ts(tsSM1_070809weekly)
```
####7.7 Create TS object with SubMeter2 - weekly (Saturday at 2pm)
```{r}
house070809weekly <- filter(newDF2, weekdays == 7 & hour == 14 & minute == 1)

tsSM2_070809weekly <- ts(house070809weekly$Sub_metering_2, frequency=52, start=c(2007,1))
```
## 7.8 Plot sub-meter 2 with autoplot - weekly
```{r}
autoplot(tsSM2_070809weekly)

## Plot sub-meter 2 with autoplot - add labels, color
autoplot(tsSM2_070809weekly, ts.colour = 'red', xlab = "Time", ylab = "Watt Hours", main = "Sub-meter 1")

## Plot sub-meter 2 with plot.ts
plot.ts(tsSM1_070809weekly)
```
#### 8. FORECASTING
####8.1 Apply time series linear regression to the sub-meter 3 ts object and use summary to obtain R2 and RMSE from the model you built
```{r}
fitSM3 <- tslm(tsSM3_070809weekly ~ trend + season) 
summary(fitSM3)

```
####8.2 Create the forecast for sub-meter 3. Forecast ahead 20 time periods 
```{r}
forecastfitSM3 <- forecast(fitSM3, h=20)
```
#### 8.3 Plot the forecast for sub-meter 3.

What do the different gray areas represent? Confidence interval
Should the forecast really have negative values, or is that an artifact from linear regression? 
```{r}
plot(forecastfitSM3)

```
####8.4 Create sub-meter 3 forecast with confidence levels 80 and 90
```{r}
forecastfitSM3c <- forecast(fitSM3, h=20, level=c(80,90))

```
## 8.5 Plot sub-meter 3 forecast, limit y and add labels
```{r}
plot(forecastfitSM3c, ylim = c(0, 20), ylab= "Watt-Hours", xlab="Time")
```
####8.6 Sub-meter 1 - linear regression model
Apply time series linear regression to the sub-meter 1 ts object and use summary to obtain R2 and RMSE from the model you built
```{r}
fitSM1 <- tslm(tsSM1_070809weekly ~ trend + season) 
summary(fitSM1)
```
####8.7 Create the forecast and plot
```{r}
forecastfitSM1 <- forecast(fitSM1, h=20)
plot(forecastfitSM1)
```
####8.8 Create sub-meter 1 forecast with confidence levels 80 and 90 - and plot
```{r}
forecastfitSM1c <- forecast(fitSM1, h=20, level=c(80,90))

plot(forecastfitSM1c, ylim = c(0, 50), ylab= "Watt-Hours", xlab="Time")
```

####8.9 Sub-meter 2 - linear regression model
Apply time series linear regression to the sub-meter 2 ts object and use summary to obtain R2 and RMSE from the model you built
```{r}
fitSM2 <- tslm(tsSM2_070809weekly ~ trend + season) 
summary(fitSM2)
```
####8.10 Create the forecast and plot
```{r}
forecastfitSM2 <- forecast(fitSM2, h=20)
plot(forecastfitSM2)
```
####8.11 Create sub-meter 2 forecast with confidence levels 80 and 90 - and plot
```{r}
forecastfitSM2c <- forecast(fitSM2, h=20, level=c(80,90))

plot(forecastfitSM2c, ylim = c(0, 80), ylab= "Watt-Hours", xlab="Time")
```
#####8.12 One comparison chart showing the R2 and RMSE of each model you built*********


#### 9. DECOMPOSING
####9.1 Decompose Sub-meter 3 into trend, seasonal and remainder
```{r}
components070809SM3weekly <- decompose(tsSM3_070809weekly)

```
####9.2 Plot decomposed sub-meter 3 
```{r}
plot(components070809SM3weekly)
```
####9.3 Check summary statistics for decomposed sub-meter 3 
```{r}
summary(components070809SM3weekly)

```
Review your resulting plot. Does sub-meter 3 show a trend in power usage? Would this information be important to a homeowner trying to understand their power consumption? Does sub-meter 3 show seasonal effects on power usage? What may or may not cause this? 
####9.4 Sub-meter 1 decomposed
```{r}
#### Decompose Sub-meter 1 into trend, seasonal and remainder

components070809SM1weekly <- decompose(tsSM1_070809weekly)

plot(components070809SM1weekly)

summary(components070809SM1weekly)

```
####9.5 Sub-meter 2 decomposed
```{r}
#### Decompose Sub-meter 2 into trend, seasonal and remainder

components070809SM2weekly <- decompose(tsSM2_070809weekly)

plot(components070809SM2weekly)

summary(components070809SM2weekly)

```
?????One comparison chart showing the summary statistics for the seasonal, trend and remainder components from each decomposed object?????
####10.HOLT WINTERS FORECASTING
####10.1 Remove seasonal components
```{r}
## Seasonal adjusting sub-meter 3 by subtracting the seasonal component & plot
tsSM3_070809Adjusted <- tsSM3_070809weekly - components070809SM3weekly$seasonal
autoplot(tsSM3_070809Adjusted)
```
## 10.2 Test Seasonal Adjustment by running Decompose again. Note the very, very small scale for Seasonal
```{r}
plot(decompose(tsSM3_070809Adjusted))


```
## 10.3 Holt Winters Exponential Smoothing & Plot
```{r}
tsSM3_HW070809 <- HoltWinters(tsSM3_070809Adjusted, beta=FALSE, gamma=FALSE)
plot(tsSM3_HW070809, ylim = c(0, 25))
```
????How does exponential smoothing help? Consider the outliers. Consider the information we removed when we subsetted millions of data points to 52 observations per year. 

Having created a ts object that contains exponentially smoothed data with no seasonality, let’s now use forecast again. 

####10.4 HoltWinters forecast & plot
```{r}
tsSM3_HW070809for <- forecast(tsSM3_HW070809, h=25)
plot(tsSM3_HW070809for, ylim = c(0, 20), ylab= "Watt-Hours", xlab="Time - Sub-meter 3")
```
let's change the the confidence levels and then plot only the forecasted area. Think of this just as you would when a weatherperson forecasts the weather: They don't include the preceding years, weeks and days. 

####10.5 Forecast HoltWinters with diminished confidence levels
```{r}
tsSM3_HW070809forC <- forecast(tsSM3_HW070809, h=25, level=c(10,25))
## Plot only the forecasted area
plot(tsSM3_HW070809forC, ylim = c(0, 20), ylab= "Watt-Hours", xlab="Time - Sub-meter 3", start(2010))
```
 ????Ask yourself, how is this plot different from the forecasted plot from step three of the plan of attack? Which, if any, is more useful? 
 
####10.6 Sub-meter1
```{r}
## Seasonal adjusting sub-meter 1 by subtracting the seasonal component & plot
tsSM1_070809Adjusted <- tsSM1_070809weekly - components070809SM1weekly$seasonal
autoplot(tsSM1_070809Adjusted)

plot(decompose(tsSM1_070809Adjusted))

tsSM1_HW070809 <- HoltWinters(tsSM1_070809Adjusted, beta=FALSE, gamma=FALSE)
plot(tsSM1_HW070809, ylim = c(0, 25))

tsSM1_HW070809for <- forecast(tsSM1_HW070809, h=25)
plot(tsSM1_HW070809for, ylim = c(0, 20), ylab= "Watt-Hours", xlab="Time - Sub-meter 1")

tsSM1_HW070809forC <- forecast(tsSM1_HW070809, h=25, level=c(10,25))
## Plot only the forecasted area
plot(tsSM1_HW070809forC, ylim = c(0, 20), ylab= "Watt-Hours", xlab="Time - Sub-meter 1", start(2010))
```
####10.7 Sub-meter2
```{r}
tsSM2_070809Adjusted <- tsSM2_070809weekly - components070809SM2weekly$seasonal
autoplot(tsSM2_070809Adjusted)

plot(decompose(tsSM2_070809Adjusted))

tsSM2_HW070809 <- HoltWinters(tsSM2_070809Adjusted, beta=FALSE, gamma=FALSE)
plot(tsSM2_HW070809, ylim = c(0, 25))

tsSM2_HW070809for <- forecast(tsSM2_HW070809, h=25)
plot(tsSM2_HW070809for, ylim = c(0, 20), ylab= "Watt-Hours", xlab="Time - Sub-meter 2")

tsSM2_HW070809forC <- forecast(tsSM2_HW070809, h=25, level=c(10,25))
## Plot only the forecasted area
plot(tsSM2_HW070809forC, ylim = c(0, 20), ylab= "Watt-Hours", xlab="Time - Sub-meter 2", start(2010))
```

